rbcL community comparison

### 2019-11-16 code/rbcl_community_compare.Rmd
updates... only using the NMDS UniFrac and the Diversity scatterplots from this script at this point, the final heatmap migrated into the community.Rmd since im comparing 18s and cellct in that one

rbcl.tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc
```

@@

```{r unifrac_plot, echo=F}
# plot
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id")

# plot unifrac, colored by elevation
uf_plot <- ggplot(unifrac_plot_data, aes(x=NMDS2, y=NMDS1)) +
  geom_point(aes(color=elev_m), size=4) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(color="Elevation (m)", tag="A") +
  geom_text_repel(aes(label=alias))
uf_plot
```

@@

otu_rel_abund_round_to_1_percent <- otu_rel_abund %>% 
  mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot

# check rel abund sums ~= 1 (will be slightly higher, due to rounding)
otu_rel_abund_round_to_1_percent %>%
  group_by(alias) %>%
  summarise(sum(rel_abund))
```

```{r plot_barplot, echo=F}
cust_theme <- theme(legend.position = "none",
                    panel.grid.major.x = element_blank(), # remove vertical lines in background
                    panel.grid.minor.x = element_blank(),
                    axis.line=element_blank(), # remove x axis
                    axis.title.x=element_blank(), 
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank(),
                    panel.spacing = unit(-0.01, "lines"), # remove space between facet panels
                    strip.text.x = element_text(angle = 90, hjust=1.0, vjust = 0.8), # rotate facet labels 90
                    strip.background = element_blank()) # remove grey boxes around facet labs

otus %>% distinct(otu_name)

otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D", "Algae E","Chloromonas F", "Trebouxiophyceae G")

barplot <- ggplot(otu_rel_abund_round_to_1_percent, aes(x=alias %>% fct_reorder(elev_m), y=rel_abund)) +
  geom_bar(aes(fill=otu_name %>% fct_relevel(otu_ord)), stat="identity") + # classic stacked relative abundance plot
  coord_flip() +
  facet_grid(~otu_name %>% fct_relevel(otu_ord), switch = "x") +
  labs( x="Sample ID", tag="B") +
  cust_theme
barplot

@@

```{r combine, fig.width=7, fig.height=10}
combined_fig_3 <- gridExtra::grid.arrange(uf_plot, barplot, nrow=2)
```

@@

```{r output, include=F}
ggsave(filename = here("manuscript/figs/3_community/combined_fig_3.png"), plot=combined_fig_3, height=10, width=7)
ggsave(filename = here("manuscript/figs/diversity.png"), plot = diversity_plots, height = 10, width = 7)
```

>>>>

rbcl.tree <- read.tree(here("output/tree/rbcl/all_asvs/rbcl_aligns_all.fasta.treefile")) ### this tree should only contain the asvs we found in the "main" sites, not looking at controls, etc

# the cell counts and 18S rel abund data
cellct_heatmap <- readRDS(here("output/cellct_w_seq_heatmap.Rda"))
s18_heatmap <- readRDS(here("output/s18_heatmap.Rda"))
```

@@

```{r unifrac_plot, echo=F, fig.width=5, fig.height=4}
# plot
unifrac_plot_data <- nmds %>% 
  scores() %>% 
  as_tibble(rownames = "sample_id") %>% 
  left_join(field, by="sample_id")

# plot unifrac, colored by elevation
uf_plot <- ggplot(unifrac_plot_data, aes(x=NMDS2, y=NMDS1)) +
  geom_point(aes(color=elev_m), size=4) +
  scale_color_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(color="Elevation (m)", tag="A") +
  geom_text_repel(aes(label=alias), size=3, segment.size = 0.3) +
  theme_classic()
uf_plot
```

@@

# otu_rel_abund_round_to_1_percent <- otu_rel_abund %>% 
#   mutate(rel_abund = if_else(rel_abund <0.01, 0.01, rel_abund)) # artificially boost rel abund <1% to 1% so its visible on plot
# 
# # check rel abund sums ~= 1 (will be slightly higher, due to rounding)
# otu_rel_abund_round_to_1_percent %>%
#   group_by(alias) %>%
#   summarise(sum(rel_abund))
```

```{r plot_barplot, echo=F, fig.width=2.9, fig.height=5}
otu_ord <- c("Chlainomonas A","Sanguina B", "Other Chloromonas","Chloromonas krienitzii","Chloromonas muramotoi C","Chloromonas D", "Unknown E","Chloromonas F", "Trebouxiophyceae G")

my_theme <- theme(axis.text.x = element_text(angle = 45, hjust = 1))

otu_rel_abund %>% distinct(sample_id) %>% nrow()

rbcl_heatmap <- ggplot(otu_rel_abund, aes(x=otu_name %>% fct_relevel(otu_ord), y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(alpha=rel_abund, fill=otu_name %>% fct_relevel(otu_ord))) +
  labs(x="rbcL OTU", y= "Elevation", fill = "rbcL OTU", alpha="Relative abundance") +
  scale_alpha(range = c(0.01, 1)) +
  my_theme +
  guides(alpha="none", fill="none")
rbcl_heatmap
```

```{r colors, include=F}
# get colors, so they match across plots
g <- ggplot_build(rbcl_heatmap)

colors <- tibble(otu_name = otu_rel_abund$otu_name,
                 color = g$data[[1]]$fill) %>% 
  distinct(otu_name, color) %>% 
  deframe()

length(colors)
colors

@@

Now add in the 18s and cellcount data side by side the rbcL OTU relative abundance data
- set colors same
- add in missing cellcount data
- plot left to right (entire width of page): rbcl, cellct, 18s, with rbcl taking up most width
- 
```{r 18s_rel_abund, fig.width=1.7, fig.height=5}
s18_heat <- s18_heatmap +
  scale_fill_manual(values=c("#D39200", "#93AA00", "#FF61C3","#808080"))
s18_heat
```

```{r cellct, fig.width=2.6, fig.height=5}
cellct_heat <- cellct_heatmap +
  scale_fill_manual(values = c("#F8766D","#D39200","#00BA38","#93AA00","#93AA00","#808080","#808080","#808080"))
cellct_heat
```

```{r elev_strip, fig.width=1.5, fig.height=5}
my_theme <- theme_classic() +
  theme(axis.line=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

elev_strip_tbl <- otu_rel_abund %>% 
  distinct(alias, elev_m) %>% 
  arrange(-elev_m) %>% 
  add_column(x="column_1") %>% 
  mutate(elev_m = elev_m %>% round(digits=-1))
nrow(elev_strip_tbl)

elev_strip <- ggplot(elev_strip_tbl, aes(x=x, y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(fill=elev_m)) +
  scale_fill_gradientn(colours = terrain.colors(8)[1:7]) +
  labs(x="Elevation (m)") +
  my_theme 
elev_strip
```

```{r elev_strip2, fig.width=0.5, fig.height=5}
elev_strip2 <- elev_strip +
  guides(fill="none") +
  geom_text(aes(label=elev_m), size=3)
elev_strip2
```

```{r snow_color_strip, fig.width=0.5, fig.height=5}
color_strip_tbl <- otu_rel_abund %>% 
  arrange(-elev_m) %>% 
  distinct(alias, color_snow, elev_m) %>% 
  add_column(x="column_1")

color_strip_tbl %>% distinct(color_snow)

color_strip <- ggplot(color_strip_tbl, aes(x=x, y=alias %>% fct_reorder(elev_m))) +
  geom_tile(aes(fill=color_snow)) +
  my_theme +
  guides(fill="none") +
  scale_fill_manual(values=c(pink="red",other="grey",green="green",orange="orange",rusty="purple"))
color_strip
```

@@

```{r output, include=F}

ggsave(here("manuscript/figs/main/community/rbcl.pdf"),plot=rbcl_heatmap, width = 2.9, height = 5)
ggsave(here("manuscript/figs/main/community/18s.pdf"),plot=s18_heat, width = 1.7, height = 5)
ggsave(here("manuscript/figs/main/community/cellct.pdf"),plot=cellct_heat, width = 2.6, height = 5)
ggsave(here("manuscript/figs/main/community/elev.pdf"),plot=elev_strip, width = 1.5, height = 5)
ggsave(here("manuscript/figs/main/community/elev2.pdf"),plot=elev_strip2, width = 0.5, height = 5)
ggsave(here("manuscript/figs/main/community/snow_color.pdf"),plot=color_strip, width = 0.5, height = 5)

ggsave(filename = here("manuscript/figs/main/uf_plot.png"), plot=uf_plot, width=5, height = 4)
ggsave(filename = here("manuscript/figs/diversity.png"), plot = diversity_plots, height = 10, width = 7)
```
